---
title: "Prediction Assignment"
author: "Ranjit Subudhi"
date: "6/23/2019"
output:
  html_document: default
  word_document: default
---
#### Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### Data Loading

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

```{r}
setwd('C:/Users/h193736/Desktop/Online Course/Data Science/7. Practical Machine Learning/Week 4')
```
### Download Packages

```{r}
library(caret)
```  

```{r}
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
```  

```{r}
library(randomForest)
```
### Data Loading and Cleaning
```{r}
Train <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Test  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```  

```{r}
training <- read.csv(url(Train))
testing  <- read.csv(url(Test))

```  
# create a partition with the training dataset 

```{r}
inTrain  <- createDataPartition(training$classe, p=0.8, list=FALSE)
TrainingSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainingSet)
```  
```{r}
dim(TestSet)
```  
# Remove Zero Variance variables 

```{r}
NZV <- nearZeroVar(TrainingSet)
TrainingSet <- TrainingSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TrainingSet)
```
```{r}
dim(TestSet)
``` 

# Remove NA variables 

```{r}
AllNA    <- sapply(TrainingSet, function(x) mean(is.na(x))) > 0.95
TrainingSet <- TrainingSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TrainingSet)
``` 

```{r}
dim(TestSet)
``` 

# remove identification only variables (columns 1 to 5)

```{r}
TrainingSet <- TrainingSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainingSet)
```

```{r}
dim(TestSet)
``` 

### Correlation Analysis

```{r}
library(corrplot)
```

```{r}
corMatrix <- cor(TrainingSet[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "lower", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))
```

###Prediction Model Building

## Random Forest

```{r}
set.seed(12345)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainingSet, method="rf",
                          trControl=controlRF)
modFitRandForest$finalModel
```


```{r}
predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest
```

```{r}
plot(confMatRandForest$table, col = confMatRandForest$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMatRandForest$overall['Accuracy'], 4)))
```

## Decision Tree

```{r}
set.seed(12345)
modFitDecTree <- rpart(classe ~ ., data=TrainingSet, method="class")
fancyRpartPlot(modFitDecTree)
```
```{r}
predictDecTree <- predict(modFitDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree
```

```{r}
plot(confMatDecTree$table, col = confMatDecTree$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confMatDecTree$overall['Accuracy'], 4)))
```

## Generalized Boosted Model

```{r}
set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainingSet, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel
```
```{r}
predictGBM <- predict(modFitGBM, newdata=TestSet)
confMatGBM <- confusionMatrix(predictGBM, TestSet$classe)
confMatGBM
```

```{r}
plot(confMatGBM$table, col = confMatGBM$byClass, 
     main = paste("GBM - Accuracy =", round(confMatGBM$overall['Accuracy'], 4)))
```


## Apply the Select Model to the Test Data
```{r}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```
